\pkg{DNest4}: Diffusive Nested Sampling in \proglang{C++} and \proglang{Python}

### 2016-05-10 paper/ms.tex
enumerate with small roman numerals
Table

\documentclass[article]{jss}
\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{dsfont}

@@

Running \pkg{DNest4} on this problem requires a class, which we call
{\tt StraightLine}, which has the following member functions:
\begin{itemize}
\item \code{void from\_prior(DNest4::RNG\& rng)}, which generates parameter
        values from the prior;
\item \code{double perturb(DNest4::RNG\& rng)}, which proposes a
        change to the parameter values, and returns $\ln(H)$ as defined in
        Equation~\ref{eqn:log_hastings};
\item \code{double log\_likelihood() const}, which evaluates the log of
        the likelihood function;
\item \code{void print(std::ostream\& out) const}, which prints parameters of
        interest to the given output stream; and
\item \code{std::string description() const}, which returns a \proglang{C++}
        string naming the parameters printed by \code{print(std::ostream\&) const}.
\end{itemize}

@@

   // Log-uniform prior
   sigma = exp(-10.0 + 20.0*rng.rand());

   // Compute the model line
   calculate_mu();
}
\end{CodeInput}
\end{CodeChunk}

@@

	// Calculate mu again if m or b changed
	if(which == 0 || which == 1)
		calculate_mu();

@@

\begin{table}[ht!]
\centering
\small
\begin{tabular}{lll}
\hline
Bayesian computation		&		Statistical mechanics		&		ABC\\
\hline
Parameter space	$\Theta$	&		Phase space	$\Omega$ or configuration space $\Gamma$ 			& (Parameter$\times$Data) space $\Theta \times \mathcal{D}$\\		
Marginal likelihood / evidence	&	Partition function at $T=1$	&\\

\end{tabular}
\caption{\it The relationship between standard Bayesian computation, statistical
mechanics, and ABC. In each case Monte Carlo methods are used to calculate
integrals over a space. In standard Bayesian computation, it is the parameter
space of a model that is usually of interest, whereas in ABC it is the space
of possible parameter values {\it and} data sets.
\label{tab:relation}}
\end{table}

@@

The main challenges associated with ABC are:
\begin{enumerate}
\item The choice of summary statistics
\item The choice of distance function
\item How to choose the value of $\epsilon$, or how to change it as a run
progresses
\item How to achieve good mixing.
\end{enumerate}
Challenges 1 and 2 are Bayesian in nature, i.e. they relate to the very
definition of the posterior distribution and are well defined independent of
the fact that we are going to use a Monte Carlo method to compute the
results. On the other hand, challenges 3 and 4 are about the computational implementation.

>>>>

\documentclass[article]{jss}
\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{enumerate}
\usepackage{dsfont}

@@

Running \pkg{DNest4} on this problem requires a class, which we call
{\tt StraightLine}, which has the following member functions:
\begin{enumerate}[(i)]
\item \code{void from\_prior(DNest4::RNG\& rng)}, which generates parameter
        values from the prior;
\item \code{double perturb(DNest4::RNG\& rng)}, which proposes a
        change to the parameter values, and returns $\ln(H)$ as defined in
        Equation~\ref{eqn:log_hastings};
\item \code{double log\_likelihood() const}, which evaluates the log of
        the likelihood function;
\item \code{void print(std::ostream\& out) const}, which prints parameters of
        interest to the given output stream; and
\item \code{std::string description() const}, which returns a \proglang{C++}
        string naming the parameters printed by \code{print(std::ostream\&) const}.
\end{enumerate}

@@

   // Log-uniform prior
   sigma = exp(-10.0 + 20.0*rng.rand());
}
\end{CodeInput}
\end{CodeChunk}

@@

@@

\begin{table}[ht!]
\centering
\small
\begin{tabular}{|l|l|l|}
\hline
Bayesian computation		&		Statistical mechanics		&		ABC\\
\hline
Parameter space	$\Theta$	&		Phase space	$\Omega$ 			& Product space $\Theta \times \mathcal{D}$\\
Posterior $p(\params | D,I)$   &  Canonical distribution $p(\boldsymbol{x}; T)$     & \\ 
Marginal likelihood $P(D|I)$	&	Partition function $Z(T)$	&   Marginal likelihood $p(D\in | I)$\\
\hline
\end{tabular}
\caption{\it The relationship between standard Bayesian computation, statistical
mechanics, and ABC. In each case Monte Carlo methods are used to calculate
integrals over a space. In standard Bayesian computation, it is the parameter
space of a model that is usually of interest, whereas in ABC it is the space
of possible parameter values {\it and} data sets.
\label{tab:relation}}
\end{table}

@@

The main challenges associated with ABC are:
\begin{enumerate}[(a)]
\item The choice of summary statistics
\item The choice of distance function
\item How to choose the value of $\epsilon$, or how to change it as a run
progresses
\item How to achieve good mixing.
\end{enumerate}
Challenges 1 and 2 are Bayesian in nature, i.e. they relate to the very
definition of the posterior distribution and are well defined independent of
the fact that we are going to use a Monte Carlo method to compute the
results. On the other hand, challenges 3 and 4 are about the computational implementation.
